Plug-and-play dual-tree algorithm runtime analysis

### 2015-10-22 bounds-all.tex
Last modifications of plug-and-play bounds paper.
Minor fixes + Dong's address update

\author{\name Ryan R. Curtin \email ryan@ratml.org\\
  \addr School of Computational Science and Engineering\\
  Georgia Institute of Technology\\
  Atlanta, GA 30332-0250, USA
  \AND
  \name Dongryeol Lee \email drselee@gmail.com\\
  \addr GE Global Research Center\\
  Schenectady, NY 12309
  \AND
  \name William B. March \email march@ices.utexas.edu\\
  \addr Institute for Computational Engineering and Sciences\\
  University of Texas, Austin\\
  Austin, TX 78712-1229
  \AND
  \name Parikshit Ram \email p.ram@gatech.edu\\
  \addr Skytree, Inc.\\
  Atlanta, GA 30332
}

@@

@@

Some of these algorithms are derived using the cover tree \citep{langford2006}, a
data structure with compelling theoretical qualities.  When cover trees are
used, Dual-tree all-nearest-neighbor search and approximate kernel density
estimation have $O(N)$ runtime guarantees for $O(N)$ queries \citep{ram2009};
minimum spanning tree calculation scales as $O(N \log N)$
\citep{march2010euclidean}.  Other problems have similar worst-case guarantees
\citep{curtin2014dual, march2013multi}.

>>>>

\author{\name Ryan R. Curtin \email ryan@ratml.org\\
  \addr School of Computational Science and Engineering\\
  Georgia Institute of Technology\\
  Atlanta, GA 30332-0250, USA
  \AND
  \name Dongryeol Lee \email drselee@gmail.com\\
  \addr Yahoo Labs\\
  Sunnyvale, CA 94089
  \AND
  \name William B. March \email march@ices.utexas.edu\\
  \addr Institute for Computational Engineering and Sciences\\
  University of Texas, Austin\\
  Austin, TX 78712-1229
  \AND
  \name Parikshit Ram \email p.ram@gatech.edu\\
  \addr Skytree, Inc.\\
  Atlanta, GA 30332
}

@@

\begin{keywords}
  dual-tree algorithms, branch and bound, nearest neighbor search, kernel
density estimation, range search
\end{keywords}

@@

Some of these algorithms are derived using the cover tree \citep{langford2006}, a
data structure with compelling theoretical qualities.  When cover trees are
used, dual-tree all-nearest-neighbor search and approximate kernel density
estimation have $O(N)$ runtime guarantees for $O(N)$ queries \citep{ram2009};
minimum spanning tree calculation scales as $O(N \log N)$
\citep{march2010euclidean}.  Other problems have similar worst-case guarantees
\citep{curtin2014dual, march2013multi}.

### 2015-10-22 bounds-files/general-runtime-bound.tex
Fix typo.
Minor fixes + Dong's address update

    \medskip
    \STATE $s^{\max}_r \gets \max_{\mathscr{N}_r \in R} s_r$
\label{alg:line:ct-dual-srmax}
    \IF{$(s_q < s^{\max}_r)$} \label{alg:line:ct-dual-ref-recursion-start}
      \STATE \COMMENT{Perform a reference recursion.}
      \FORALL{$\mathscr{N}_r \in R$} \label{alg:line:ct-dual-base-case-start}
        \STATE \texttt{BaseCase($p_q$, $p_r$)}
      \ENDFOR \label{alg:line:ct-dual-base-case-end}
      \STATE $R_r \gets \{ \mathscr{N}_r \in R : s_r = s^{\max}_r \}$
\label{alg:line:ct-dual-ref-set}
      \STATE $R_{r - 1} \gets \{ \mathscr{C}(\mathscr{N}_r) : \mathscr{N}_r \in
R_r \} \cup (R \setminus R_r)$ \label{alg:line:ct-dual-ref-children}
      \STATE $R'_{r - 1} \gets \{ \mathscr{N}_r \in R_{r - 1} :
\texttt{Score(}\mathscr{N}_q\texttt{,} \mathscr{N}_r\texttt{)} \ne \infty \}$
\label{alg:line:ct-dual-ref-score}
      \STATE recurse with $\mathscr{N}_q$ and $R'_{r - 1}$
\label{alg:line:ct-dual-ref-recursion-end}
    \ELSE \label{alg:line:ct-dual-query-recursion-start}
      \STATE \COMMENT{Perform a query recursion.}
      \FORALL{$\mathscr{N}_{qc} \in \mathscr{C}(\mathscr{N}_q)$}
        \STATE $R' \gets \{ \mathscr{N}_r \in R :
\texttt{Score(}\mathscr{N}_q\texttt{,} \mathscr{N}_r\texttt{)} \ne \infty \}$
\label{alg:line:ct-dual-query-pruning}
        \STATE recurse with $\mathscr{N}_{qc}$ and $R'$
\label{alg:line:ct-dual-query-recursion}
      \ENDFOR \label{alg:line:ct-dual-query-recursion-end}
    \ENDIF
  \end{algorithmic}
  \caption{The standard pruning dual-tree traversal for cover trees.}
  \label{alg:cover-tree-dual}
\end{algorithm}

@@

Now suppose that $s_i < s_r^{\max} - 1$.  This implies that there exists some
implicit cover tree node with point $p_i$ and scale $s_r^{\max} - 1$ (as well an
implicit child of this node $p_i$ with scale $s_r^{\max} - 2$ and so forth until
one of these implicit nodes has child $p_i$ with scale $s_i$).  Because the
separation invariant applies to both implicit and explicit representations of
the tree, we conclude that $d(p_i, p_r) > 2^{s_r^{\max}} - 1$.  The same
argument may be made for the case where $s_j < s_r^{\max} - 1$, with the same
conclusion.

>>>>

    \medskip
    \STATE $s^{\max}_r \gets \max_{\mathscr{N}_r \in R} s_r$
\label{alg:line:ct-dual-srmax}
    \IF{$(s_q < s^{\max}_r)$} \label{alg:line:ct-dual-ref-recursion-start}
      \STATE \COMMENT{Perform a reference recursion.}
      \FORALL{$\mathscr{N}_r \in R$} \label{alg:line:ct-dual-base-case-start}
        \STATE \texttt{BaseCase($p_q$, $p_r$)}
      \ENDFOR \label{alg:line:ct-dual-base-case-end}
      \STATE $R_r \gets \{ \mathscr{N}_r \in R : s_r = s^{\max}_r \}$
\label{alg:line:ct-dual-ref-set}
      \STATE $R_{r - 1} \gets \{ \mathscr{C}(\mathscr{N}_r) : \mathscr{N}_r \in
R_r \} \cup (R \setminus R_r)$ \label{alg:line:ct-dual-ref-children}
      \STATE $R'_{r - 1} \gets \{ \mathscr{N}_r \in R_{r - 1} :
\texttt{Score(}\mathscr{N}_q\texttt{,} \mathscr{N}_r\texttt{)} \ne \infty \}$
\label{alg:line:ct-dual-ref-score}
      \STATE recurse with $\mathscr{N}_q$ and $R'_{r - 1}$
\label{alg:line:ct-dual-ref-recursion-end}
    \ELSE \label{alg:line:ct-dual-query-recursion-start}
      \STATE \COMMENT{Perform a query recursion.}
      \FORALL{$\mathscr{N}_{qc} \in \mathscr{C}(\mathscr{N}_q)$}
        \STATE $R' \gets \{ \mathscr{N}_r \in R :
\texttt{Score(}\mathscr{N}_{qc}\texttt{,} \mathscr{N}_r\texttt{)} \ne \infty \}$
\label{alg:line:ct-dual-query-pruning}
        \STATE recurse with $\mathscr{N}_{qc}$ and $R'$
\label{alg:line:ct-dual-query-recursion}
      \ENDFOR \label{alg:line:ct-dual-query-recursion-end}
    \ENDIF
  \end{algorithmic}
  \caption{The standard pruning dual-tree traversal for cover trees.}
  \label{alg:cover-tree-dual}
\end{algorithm}

@@

Now suppose that $s_i < s_r^{\max} - 1$.  This implies that there exists some
implicit cover tree node with point $p_i$ and scale $s_r^{\max} - 1$ (as well an
implicit child of this node $p_i$ with scale $s_r^{\max} - 2$ and so forth until
one of these implicit nodes has child $p_i$ with scale $s_i$).  Because the
separation invariant applies to both implicit and explicit representations of
the tree, we conclude that $d(p_i, p_j) > 2^{s_r^{\max}} - 1$.  The same
argument may be made for the case where $s_j < s_r^{\max} - 1$, with the same
conclusion.

### 2015-10-22 bounds-files/sec01-supporting-lemmas.tex
Minor fixes + Dong's address update

\begin{table}
{\small
\begin{center}
\begin{tabular}{|c|l|}
\hline
{\bf Symbol} & {\bf Description} \\ \hline
$\mathscr{N}$ & A tree node \\ \hline
$\mathscr{C}_i$ & Set of child nodes of $\mathscr{N}_i$ \\ \hline
$\mathscr{P}_i$ & Set of points held in $\mathscr{N}_i$ \\ \hline
$\mathscr{D}_i^n$ & Set of descendant nodes of $\mathscr{N}_i$ \\ \hline
$\mathscr{D}_i^p$ & Set of points contained in $\mathscr{N}_i$ and
$\mathscr{D}_i^n$ \\ \hline
$\mu_i$ & Center of $\mathscr{N}_i$ \\ \hline
$\lambda_i$ & Furthest descendant distance \\ \hline
\end{tabular}
\end{center}
}
\caption{Notation for trees.  See \cite{curtin2013tree} for details.}
\label{tab:notation}
\end{table}

>>>>

\begin{table}
{\small
\begin{center}
\begin{tabular}{|c|l|}
\hline
{\bf Symbol} & {\bf Description} \\ \hline
$\mathscr{N}$ & A tree node \\ \hline
$\mathscr{C}_i$ & Set of child nodes of $\mathscr{N}_i$ \\ \hline
$\mathscr{P}_i$ & Set of points held in $\mathscr{N}_i$ \\ \hline
$\mathscr{D}_i^n$ & Set of descendant nodes of $\mathscr{N}_i$ \\ \hline
$\mathscr{D}_i^p$ & Set of points contained in $\mathscr{N}_i$ and
$\mathscr{D}_i^n$ \\ \hline
$\mu_i$ & Center of $\mathscr{N}_i$ \\ \hline
$\lambda_i$ & Furthest descendant distance from $\mu_i$\\ \hline
\end{tabular}
\end{center}
}
\caption{Notation for trees.  See \cite{curtin2013tree} for details.}
\label{tab:notation}
\end{table}
